---
title: "Doxycyline effect"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = FALSE,
        dplyr.summarise.inform = FALSE,
        tidyverse.quiet = TRUE)
library(tidyverse)
library(DESeq2)

# install.packages("seriation")
library(seriation)
source("../../util/_plot_theme.R")
source("../../util/_util.R")
source("../../util/deseq.R")
source("../../util/figures.R")
source("../01_setup/assumptions.R")

# Keep track of thresholds for figure naming
thresh <- paste0("pval", pval_thresh, "_l2fc", round(l2fc_thresh, 2))

# we have set our standard thresholds to :
# P <0.05 & Lfc .58
```

### First we load in the KO control line (transgene lacking rTTA ) over the longt time course

```{r loading in DeSeq results for esc_ko_control_long time course}

#TODO How is this file derived (looks like Deseq format --- )
#TODO better understand "esc_ko_control_long.RData 
if(!file.exists("results/esc_ko_control_long.RData")) {
  source("../../util/deseq.R")
  esc_ko_control_long <- deseq_vs_zero(ct = "ESC", 
                                       tc_len = "long", 
                                       fko = "KO", 
                                       fi = "control")
  save(esc_ko_control_long, file = "results/esc_ko_control_long.RData")
}

load("results/esc_ko_control_long.RData", verbose = T)
```

# Filtering to significant and lfc thresholds
```{r P < 0.05 and lfc > 0.58}

# Filtering esc_ko_control long
# mutate to make a col of sig if meets thresholds
esc_ko_control_long <- esc_ko_control_long %>%
  mutate(sig = padj <= pval_thresh & max_abs_l2fc_shrunken > l2fc_thresh)

# creating sig gene object for dox effects
# filtering on new sig col from above
esc_ko_control_long_sig <- esc_ko_control_long %>% 
  filter(sig == TRUE)

```

# Plotting sig genes across time using "make_timecourse_lfc_plot
```{r Plotting sig genes across time using "make_timecourse_lfc_plot}

# Plotting
make_timecourse_lfc_plot(deseq_res = esc_ko_control_long,
                         genes_to_include = unique(esc_ko_control_long_sig$gene_id),
                         y_lim = c(-5.5,8.2),
                         draw_plot = FALSE)
ggsave(paste0("figures/esc_ko_control_long_lfc_", thresh, ".pdf"), 
       height = 3, width = 3.5, useDingbats = FALSE)
```
### Result: Lots of genes are affected by Dox after 12 hours -- not much before
### TERC one of the most down regulated (~20x) and H19 one most upregulated (2~0x)


# Dox affected genes in KO background contorl short time course
# The first time run - Deseq will be invoked to differential analysis
# Per the "deseq.R" function "deseq_vs_zero"
```{r KO control short time course dox genes}


# running Deseq on sample characteristics selected in "deseq_vs_zero"
if(!file.exists("results/esc_ko_control_short.RData")) {
  source("../util/deseq.R")
  esc_ko_control_short <- deseq_vs_zero(ct = "ESC", 
                                       tc_len = "short", 
                                       fko = "KO", 
                                       fi = "control")
  save(esc_ko_control_short, file = "results/esc_ko_control_short.RData")
}

load("results/esc_ko_control_short.RData", verbose = T)
```


```{r KO control short time course dox genes P < 0.05 and lfc > 0.58 }
esc_ko_control_short <- esc_ko_control_short %>%
  mutate(sig = padj <= pval_thresh & max_abs_l2fc_shrunken > l2fc_thresh)

esc_ko_control_short_sig <- esc_ko_control_short %>% 
  filter(sig == TRUE)
```

# Plotting KO control short time course dox genes
```{r Figure : KO control short time course }
make_timecourse_lfc_plot(deseq_res = esc_ko_control_short,
                         genes_to_include = unique(esc_ko_control_short_sig$gene_id),
                         draw_plot = FALSE)

# ggsave(paste0("figures/esc_ko_control_short_lfc_", thresh, ".pdf"), 
#        height = 3, width = 3.5, useDingbats = FALSE)
```
### Result: many fewer genes are changing -- Neat 1 is going up


# WT control long time course Deseq run = Dox genes
```{r WT control long time course Deseq run}

# setting Deseq parameters and running for WT control long time course
if(!file.exists("results/esc_wt_control_long.RData")) {
  source("../util/deseq.R")
  esc_wt_control_long <- deseq_vs_zero(ct = "ESC", 
                                       tc_len = "long", 
                                       fko = "WT", 
                                       fi = "control")
  save(esc_wt_control_long, file = "results/esc_wt_control_long.RData")
}

load("results/esc_wt_control_long.RData", verbose = T)
```

# Selecting for significant genes in WT control long time course
```{r WT control dox genes P < 0.05 and lfc > 0.58 }

esc_wt_control_long <- esc_wt_control_long %>%
  mutate(sig = padj <= pval_thresh & max_abs_l2fc_shrunken > l2fc_thresh)

esc_wt_control_long_sig <- esc_wt_control_long %>% 
  filter(sig == TRUE)
```

# Plotting WT control sig dox genes long time course
```{r Plotting WT control sig dox genes long time course}

make_timecourse_lfc_plot(deseq_res = esc_wt_control_long,
                         genes_to_include = unique(esc_wt_control_long_sig$gene_id),
                         y_lim = c(-5,5),
                         draw_plot = FALSE)

# ggsave(paste0("figures/esc_wt_control_long_lfc_", thresh, ".pdf"), 
#        height = 3, width = 3.5, useDingbats = FALSE)
```
### Result: hundreds of genes changing after 12 hours due to dox
### Result: similar to KO control long time course
### Result: Short time course does not have as many sig dox affects in WT or KO backgorund 



# Running Deseq on wt control short time course 
```{r WT control short time course Deseq }

if(!file.exists("results/esc_wt_control_short.RData")) {
  source("../util/deseq.R")
  esc_wt_control_short <- deseq_vs_zero(ct = "ESC", 
                                       tc_len = "short", 
                                       fko = "WT", 
                                       fi = "control")
  save(esc_wt_control_short, file = "results/esc_wt_control_short.RData")
}

load("results/esc_wt_control_short.RData", verbose = T)
```

# Filtering to significant genes in WT control short time course
```{r WT control short time course P < 0.05 and lfc > 0.58}

esc_wt_control_short <- esc_wt_control_short %>%
  mutate(sig = padj <= pval_thresh & max_abs_l2fc_shrunken > l2fc_thresh)

esc_wt_control_short_sig <- esc_wt_control_short %>% 
  filter(sig == TRUE)


```

# Plotting wt control short time course sig dox genes 
```{r Plotting wt control short time course sig dox genes }

make_timecourse_lfc_plot(deseq_res = esc_wt_control_short,
                         genes_to_include = unique(esc_wt_control_short_sig$gene_id),
                         draw_plot = FALSE)

# ggsave(paste0("figures/esc_wt_control_short_lfc_", thresh, ".pdf"), 
#        height = 3, width = 3.5, useDingbats = FALSE)
```
### Result: again short time course is many fewer genes due to dox effects
### Result: WT has more changes due to dox 63 vs 7 in K0 short




# Deseq on NPC KO control Short timecourse dox effect
```{r Deseq on NPC KO control Short timecourse dox effect}

if(!file.exists("results/npc_ko_control_short.RData")) {
  source("../util/deseq.R")
  npc_ko_control_short <- deseq_vs_zero(ct = "NPC", 
                                       tc_len = "short", 
                                       fko = "KO", 
                                       fi = "control")
  save(npc_ko_control_short, file = "results/npc_ko_control_short.RData")
}

load("results/npc_ko_control_short.RData", verbose = T)
```

# Filtering to signficant genes P < 0.05 and lfc > 0.58
```{r NPC KO control short time course P < 0.05 and lfc > 0.58}

npc_ko_control_short <- npc_ko_control_short %>%
  mutate(sig = padj <= pval_thresh & max_abs_l2fc_shrunken > l2fc_thresh)

npc_ko_control_short_sig <- npc_ko_control_short %>% 
  filter(sig == TRUE)
```
### Result: There are no significant genes for NPC KO control.

# Deseq on NPC WT Short timecourse dox effect

```{r NPC WT control Deseq }

if(!file.exists("results/npc_wt_control_short.RData")) {
  source("../util/deseq.R")
  npc_wt_control_short <- deseq_vs_zero(ct = "NPC", 
                                       tc_len = "short", 
                                       fko = "WT", 
                                       fi = "control")
  save(npc_wt_control_short, file = "results/npc_wt_control_short.RData")
}

load("results/npc_wt_control_short.RData", verbose = T)
```

# Filtering to significant dox genes in WT control NPCs
```{r WT control NPCs dox genes P < 0.05 and lfc > 0.58}

npc_wt_control_short <- npc_wt_control_short %>%
  mutate(sig = padj <= pval_thresh & max_abs_l2fc_shrunken > l2fc_thresh)

npc_wt_control_short_sig <- npc_wt_control_short %>% 
  filter(sig == TRUE)
```

# Plotting WT NPC control short time course dox genes
```{r lotting WT NPC control short time course dox genes}

make_timecourse_lfc_plot(deseq_res = npc_wt_control_short,
                         genes_to_include = unique(npc_wt_control_short_sig$gene_id),
                         draw_plot = FALSE)

# ggsave(paste0("figures/npc_wt_control_short_lfc_", thresh, ".pdf"), 
#        height = 3, width = 3.5, useDingbats = FALSE)
```
### Result: only one gene is upregulated due to dox


